---
title: "Reproducible Research Assignment"
author: "Syamala"
date: "January 27, 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading Acitivty Data [file should be in current working directory]
```{r}
        ActivityData <- read.csv("activity.csv", header = TRUE)
```


## Total number of steps taken per day
```{r }
        DailySteps <- aggregate(ActivityData[,1], list(sum = ActivityData$date), sum, na.rm = TRUE)
```

## Histogram of daily total steps
```{r }

        hist(DailySteps$x ,main = "Histogram of daily total steps" , xlab = "Total steps")
```

## Mean of the total number of steps taken per day
```{r }
        tapply(ActivityData$steps, ActivityData$date, mean ,na.rm = TRUE)
```

## Median of the total number of steps taken per day   
```{r}
        tapply(ActivityData$steps, ActivityData$date, median, na.rm = TRUE)
```

##Average daily activity pattern  
```{r}
        FiveMinAvg <- aggregate(ActivityData[,1], list(Interval = ActivityData$interval), mean, na.rm = TRUE)
      
        plot(x ~ Interval, FiveMinAvg , type = "l" , ylab = "Average Steps taken" )
         
```        

#Interval that contains the maximum number of steps  
```{r}
        FiveMinAvg[which.max(FiveMinAvg$x), ]

```


##Total number of missing values
```{r}
        sum(is.na(ActivityData$steps))
```

##Imputing missing values
```{r}
        require(plyr)
        require(Hmisc)
        imputeValue <- FiveMinAvg[which(FiveMinAvg$Interval == 5), ]
        ActivityData$new <- with(ActivityData, impute(steps, imputeValue$x))
        NewActivityData <- data.frame(ActivityData$new, ActivityData$date, ActivityData$interval)
        colnames(NewActivityData) <- c("steps", "date", "interval")       
```
## Total number of steps taken per day after imputing missing values  
```{r }
        NewDailySteps <- aggregate(NewActivityData[,1], list(sum = NewActivityData$date), sum, na.rm = TRUE)
```   
## Histogram of daily total steps after imputing missing values
```{r }
        hist(NewDailySteps$x, main = "Histogram of daily total steps" , xlab = "Total steps")
```  

## Mean after imputing missing values
```{r }
        tapply(NewActivityData$steps, NewActivityData$date, mean ,na.rm = TRUE)
```

## Median after imputing missing values   
```{r}
        tapply(NewActivityData$steps, NewActivityData$date, median, na.rm = TRUE)
```
##Activity patterns between weekdays and weekends
```{r}
        NewActivityData$date <- as.Date(NewActivityData$date)
       #create a vector of weekdays
       weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
  
       NewActivityData$wDay <- factor((weekdays(NewActivityData$date) %in% weekdays1),levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

     WeekDataset <- aggregate(NewActivityData[,1] , list(ActivityData$interval, NewActivityData$wDay), mean )
     colnames(WeekDataset) <- c("interval", "date", "avg") 
     xyplot(avg~interval|factor(date), data=WeekDataset,  type='l',layout=c(1,2), xlab='Interval',ylab='Number of Steps')
```
